---
title: 'Homework 1'
header-includes:
  - \usepackage{multirow}
output:
  pdf_document: default
urlcolor: blue
---

```{r, include=FALSE}

knitr::opts_chunk$set(tidy = FALSE, message = F, warning = F)
```


Github link: [https://github.com/xxou617/bios731_hw1_ou](https://github.com/xxou617/bios731_hw1_ou)

$$Y_i = \beta_0 + \beta_{treatment}X_{i1} + \mathbf{Z_i}^T\boldsymbol{\gamma} + \epsilon_i$$


Notation is defined below: 

* $Y_i$: continuous outcome
* $X_{i1}$: treatment group indicator; $X_{i1}=1$ for treated 
* $\mathbf{Z_i}$: vector of potential confounders
* $\beta_{treatment}$: average treatment effect, adjusting for $\mathbf{Z_i}$
* $\boldsymbol{\gamma}$: vector of regression coefficient values for confounders 
* $\epsilon_i$: errors, we will vary how these are defined


In our simulation, we want to 

* Estimate $\beta_{treatment}$ and $se(\hat{\beta}_{treatment})$
  * Evaluate $\beta_{treatment}$ through bias and coverage
  * We will use 3 methods to compute $se(\hat{\beta}_{treatment})$ and coverage:
    1. Wald confidence intervals (the standard approach)
    2. Nonparametric bootstrap percentile intervals
    3. Nonparametric bootstrap $t$ intervals
  * Evaluate computation times for each method to compute a confidence interval

* Evaluate these properties at:
  - Sample size $n \in \{10, 50, 500\}$
  - True values $\beta_{treatment} \in \{0, 0.5, 2\}$
  - True $\epsilon_i$ normally distributed with $\epsilon_i \sim N(0, 2)$
  - True $\epsilon_i$ coming from a right skewed distribution
    - **Hint**: try $\epsilon_i \sim logNormal(0, \log (2))$

* Assume that there are no confounders ($\boldsymbol{\gamma} = 0$)
* Use a full factorial design



### Problem 1.1 ADEMP Structure 

Answer the following questions: 

#### (1) How many simulation scenarios will you be running?

- 3 $\beta_{treatment}$
- 3 sample sizes
- 3 methods 
- 2 error term distribution

$$
3*3*2*3 = 54
$$

Thus, we have 54 scenarios in total.

#### (2) What are the estimand(s)

- $\beta_{treatment}$ 

#### (3) What method(s) are being evaluated/compared?

- Wald confidence intervals (the standard approach), nonparametric bootstrap percentile intervals and nonparametric bootstrap $t$ intervals for $se(\hat{\beta}_{treatment})$ in different scenario will be evaluated.

#### (4) What are the performance measure(s)?

- bias
- $se(\hat{\beta})$
- coverage
- computation time


### Problem 1.2 nSim 

Based on desired coverage of 95\% with Monte Carlo error of no more than 1\%, how many simulations ($n_{sim}$) should we perform for each simulation scenario?  Implement this number of simulations throughout your simulation study.

$$
n_{sim} = \frac{0.95(1-0.95)}{(0.01)^2} = 475
$$

### Problem 1.3 Implementation 


We will execute this full simulation study. For full credit, make sure to implement the following:

* Well structured scripts and subfolders following guidance from `project_organization` lecture
* Use relative file paths to access intermediate scripts and data objects
* Use readable code practices
* Parallelize your simulation scenarios
* Save results from each simulation scenario in an intermediate `.Rda` or `.rds` dataset in a `data` subfolder 
  * Ignore these data files in your `.gitignore` file so when pushing and committing to GitHub they don't get pushed to remote 
* Make sure your folder contains a Readme explaining the workflow of your simulation study
  * should include how files are executed and in what order 
* Ensure reproducibility! I should be able to clone your GitHub repo, open your `.Rproj` file, and run your simulation study



### Problem 1.4 Results summary

Create a plot or table to summarize simulation results across scenarios and methods for each of the following. 

- Bias of $\hat{\beta}$
- Coverage of $\hat{\beta}$
- Distribution of $se(\hat{\beta})$
- Computation time across methods


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(gt)
```

```{r}
load(here::here("data", "all_scenarios.Rdata"))
```

```{r}
# summarise results grouped by the scenarios

final_results <- final_results |> 
  mutate(
    beta_true = as.factor(beta_true),
    epsilon_distr = factor(epsilon_distr, levels = c("Normal", "logNormal")),
    CI_method = factor(CI_method, levels = c("Wald", "boot quantile", "boot t"))
  )

simu_evaluate <- final_results |> 
  group_by(n, beta_true, epsilon_distr, CI_method) |> 
  summarise(
    bias = mean(error),
    avg_se = mean(est_se),
    avg_time = mean(cal_CI_time),
    coverage_rate = mean(coverage),
    .groups = "drop" 
  ) 

head(simu_evaluate)
```


```{r, fig.align='center', fig.width = 7, fig.height= 6}
sample_size = unique(simu_evaluate$n)


# Plot for bias
ggplot(simu_evaluate, 
       aes(x = n, y = bias, color = beta_true, group = beta_true)) +
  geom_point(size = 2) +  
  geom_line() +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") + 
  labs(x = "Sample Size", y = "Bias", color = "Beta True",
       title = "Bias Across Sample Sizes for \nDifferent CI Methods and Error Distributions") +  
  theme_bw() +
  theme(legend.position = "top") +
  facet_grid( cols = vars(CI_method), rows = vars(epsilon_distr)) +
  scale_x_continuous(breaks = sample_size)  -> plot_bias

plot_bias
```

```{r, fig.align='center', fig.width = 7, fig.height= 6}

# plot for coverage

ggplot(simu_evaluate, 
       aes(x = n, y = coverage_rate, color = beta_true, group = beta_true)) +
  geom_point(size = 2) +  
  geom_line() +  
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "blue") +  
  labs(x = "Sample Size", y = "Coverage", color = "Beta True",
       title = "Coverage rate Across Sample Sizes for \nDifferent CI Methods and Error Distributions") +  
  theme_bw() +
  theme(legend.position = "top") +
  facet_grid( cols = vars(CI_method), rows = vars(epsilon_distr)) +
  scale_x_continuous(breaks = sample_size)  -> plot_coverage

plot_coverage

```

```{r, fig.align='center', fig.width = 6, fig.height= 6}

# distribution of se(beta_hat)
ggplot(final_results, 
       aes(x = as.factor(n), y = est_se, color = beta_true)) +
  geom_boxplot() +
  labs(x = "Sample Size", 
       y = "Estimated Standard Error", 
       color = "Beta True",
       title = "Distribution of Estimated Standard Errors Across \nSample Sizes, CI Methods, and Error Distributions") +
  theme_bw() +
  theme(legend.position = "top") +
  facet_grid(cols = vars(CI_method), rows = vars(epsilon_distr), scales = "free_y") +
  scale_x_discrete(labels = sample_size) -> plot_se

plot_se

```

```{r, fig.align='center', fig.width = 7, fig.height= 8}

# plot for calculation time
ggplot(final_results, aes(x = as.factor(n), y = cal_CI_time, color = beta_true)) +
  geom_boxplot() +
  labs(x = "Sample Size", 
       y = "Computation Time", 
       color = "Beta True",
       title = "Distribution of Time for CI Calculation Across \nSample Sizes, CI Methods, and Error Distributions") +
  theme_bw() +
  theme(legend.position = "top") +
  facet_wrap(~epsilon_distr+CI_method, scales = "free_y") +
  scale_x_discrete(labels = sample_size) -> plot_time

plot_time
``` 

```{r}

# save plots in results folder
ggsave(plot_bias, file = here::here("results", "plot_bias.jpg"),
       width = 7, height = 6)
ggsave(plot_coverage, file = here::here("results", "plot_coverage.jpg"),
       width = 7, height = 6)
ggsave(plot_se, file = here::here("results", "plot_se.jpg"),
       width = 6, height = 6)
ggsave(plot_time, file = here::here("results", "plot_time.jpg"),
       width = 7, height = 8)

```


### Problem 1.5 Discussion

Interpret the results summarized in Problem 1.4. First, write a **paragraph** summarizing the main findings of your simulation study. Then, answer the specific questions below.

The simulation study evaluated three confidence interval (CI) methods—Wald intervals, bootstrapped quantile intervals, and bootstrapped t intervals—across varying sample sizes, treatment effects, and error distributions. Wald CI was the fastest but performed poorly under skewed errors. Boot quantile CI had moderate computation time, while Boot t CI improved coverage in skewed data but required the most computational resources. As sample size increased, bias decreased under the normal distribution but increased in the log-normal distribution when the treatment effect was nonzero, suggesting that GLM with a log link may not be ideal for estimating $\beta_{treatment}$ in this setting. The estimated standard error  decreased as sample size grew, with Boot quantile CI producing the smallest standard errors. 

- How do the different methods for constructing confidence intervals compare in terms of computation time?

Wald CI is the fastest, boot quantile CI requires Moderate computation time, and boot t CI is the slowest and requires significantly more computation time than the other two methods.

- Which method(s) for constructing confidence intervals provide the best coverage when $\epsilon_i \sim N(0, 2)$?

Wald CI


- Which method(s) for constructing confidence intervals provide the best coverage when $\epsilon_i \sim logNormal(0, \log (2))$?

boot t CI




# Additional discussion


Below are the outcomes when the model is fitted using `lm()` for both the normal and log-normal distributions. In this case, the Wald CI provides the best coverage for both $\epsilon_i \sim N(0, 2)$ and $\epsilon_i \sim logNormal(0, \log (2))$


```{r}
load(here::here("data", "all_scenarios0.Rdata"))
```

```{r}
# summarise results grouped by the scenarios

final_results <- final_results |> 
  mutate(
    beta_true = as.factor(beta_true),
    epsilon_distr = factor(epsilon_distr, levels = c("Normal", "logNormal")),
    CI_method = factor(CI_method, levels = c("Wald", "boot quantile", "boot t"))
  )

simu_evaluate <- final_results |> 
  group_by(n, beta_true, epsilon_distr, CI_method) |> 
  summarise(
    bias = mean(error),
    avg_se = mean(est_se),
    avg_time = mean(cal_CI_time),
    coverage_rate = mean(coverage),
    .groups = "drop" 
  ) 

head(simu_evaluate)
```


```{r, fig.align='center', fig.width = 7, fig.height= 6}
sample_size = unique(simu_evaluate$n)


# Plot for bias
ggplot(simu_evaluate, 
       aes(x = n, y = bias, color = beta_true, group = beta_true)) +
  geom_point(size = 2) +  
  geom_line() +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") + 
  labs(x = "Sample Size", y = "Bias", color = "Beta True",
       title = "Bias Across Sample Sizes for \nDifferent CI Methods and Error Distributions") +  
  theme_bw() +
  theme(legend.position = "top") +
  facet_grid( cols = vars(CI_method), rows = vars(epsilon_distr)) +
  scale_x_continuous(breaks = sample_size)  -> plot_bias

plot_bias
```

```{r, fig.align='center', fig.width = 7, fig.height= 6}

# plot for coverage

ggplot(simu_evaluate, 
       aes(x = n, y = coverage_rate, color = beta_true, group = beta_true)) +
  geom_point(size = 2) +  
  geom_line() +  
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "blue") +  
  labs(x = "Sample Size", y = "Coverage", color = "Beta True",
       title = "Coverage rate Across Sample Sizes for \nDifferent CI Methods and Error Distributions") +  
  theme_bw() +
  theme(legend.position = "top") +
  facet_grid( cols = vars(CI_method), rows = vars(epsilon_distr)) +
  scale_x_continuous(breaks = sample_size)  -> plot_coverage

plot_coverage

```

```{r, fig.align='center', fig.width = 6, fig.height= 6}

# distribution of se(beta_hat)
ggplot(final_results, 
       aes(x = as.factor(n), y = est_se, color = beta_true)) +
  geom_boxplot() +
  labs(x = "Sample Size", 
       y = "Estimated Standard Error", 
       color = "Beta True",
       title = "Distribution of Estimated Standard Errors Across \nSample Sizes, CI Methods, and Error Distributions") +
  theme_bw() +
  theme(legend.position = "top") +
  facet_grid(cols = vars(CI_method), rows = vars(epsilon_distr), scales = "free_y") +
  scale_x_discrete(labels = sample_size) -> plot_se

plot_se

```

```{r, fig.align='center', fig.width = 7, fig.height= 8}

# plot for calculation time
ggplot(final_results, aes(x = as.factor(n), y = cal_CI_time, color = beta_true)) +
  geom_boxplot() +
  labs(x = "Sample Size", 
       y = "Computation Time", 
       color = "Beta True",
       title = "Distribution of Time for CI Calculation Across \nSample Sizes, CI Methods, and Error Distributions") +
  theme_bw() +
  theme(legend.position = "top") +
  facet_wrap(~epsilon_distr+CI_method, scales = "free_y") +
  scale_x_discrete(labels = sample_size) -> plot_time

plot_time
``` 



